---
title: "Data Cleaning lab"
author: "Angie Boysen"
date: "July 12, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr)
library(stringr)
```

### A. "Real Property Taxes" Dataset 

Download the "real property taxes" dataset from the website (via OpenBaltimore), 
he data is located here (note you don\'t need to unzip it to read it into R):
http://sisbid.github.io/Module1/data/real_property_tax.csv.gz

1. Read the Property Tax data into R and call it the variable `tax`
```{r read}
tax <- read_csv("real_property_tax.csv.gz")
str(tax)
head(tax,2)
anyNA(tax$propertyAddress)
```

2. How many addresses pay property taxes? 
```{r number}
nrow(tax[-which(is.na(tax$propertyAddress)),])
```
 
3. What is the total city and state tax paid?
```{r }
totalCity <- sum(tax$cityTax)
new_city <- tax
new_city <- new_city %>% mutate(cityTax = cityTax %>%
                                str_replace(fixed("$"),"") %>%
                                as.numeric)
totalCity <- sum(new_city$cityTax, na.rm = T)
totalCity

new_state <- new_city
new_state <- new_state %>% mutate(stateTax = stateTax %>%
                                str_replace(fixed("$"),"") %>%
                                as.numeric)
totalState <- sum(new_state$stateTax, na.rm = T)
totalState
tax <- new_state
```

The total city tax is `r totalCity`, and the total state tax is `r totalState`.

4. How many observations/properties are in each residence code?
```{r resCode}
anyNA(tax$resCode)
table(tax$resCode)
```


5. What is the 75th percentile of city and state tax paid by residence code?
```{r}
codes <- unique(tax$resCode)
# perc75 <- c()
# for(i in 1:length(codes)){
#      city <- quantile(tax$cityTax[tax$resCode == codes[i]], na.rm = T)["75%"]
#      names(city) <- paste(codes[i],"city", tax)
#      state <-quantile(tax$stateTax[tax$resCode == codes[i]], na.rm = T)["75%"]
#      names(state) <- paste(codes[i],"state", tax)
# 
#      perc75 <- c(perc75,  city, state)
# }
# perc75
ss.tax <- filter(tax,resCode == codes[1])
not.princ <- sapply(ss.tax[,str_subset(names(ss.tax),"Tax")],function(x) quantile(x,na.rm=T)["75%"])
ss.tax <- filter(tax,resCode == codes[2])
princ <- sapply(ss.tax[,str_subset(names(ss.tax),"Tax")],function(x) quantile(x,na.rm=T)["75%"])
not.princ
princ
```


6. Subset the data to only retain those houses that are principal residences and describe the distribution of property taxes on these residences.
```{r}
ss.tax <- filter(tax,resCode == codes[2])
sapply(ss.tax[,str_subset(names(ss.tax),"Tax")],quantile, na.rm=T)

```


7. Convert the 'lotSize' variable to a numeric square feet variable. Tips:
* Look at the data 
* Assume hyphens represent inches within square foot meassuremnts 
* Assume decimals within acreage measurements
* 1 acre = 43560 square feet
* Look at the data 
```{r}
acres.rows <- grep("ACRES",tax$lotSize)
new_tax <- tax
new_tax[acres.rows,] <- new_tax[acres.rows,] %>% mutate(lotSize = lotSize %>%
                                str_replace(("ACRES"),"") %>%
                                str_trim())

dash.rows <- grep("-",new_tax$lotSize,fixed = T)
```


### B. "Baltimore Salary 2015" Dataset 

Download the "Salary 2015" Dataset from the website (via OpenBaltimore), which is located here:
http://sisbid.github.io/Module1/data//Baltimore_City_Employee_Salaries_FY2015.csv
```{r}
salary <- read.csv("Baltimore_City_Employee_Salaries_FY2015.csv")
names(salary)
```

8. Make an object called health.sal using the salaries data set,	with only agencies of those with "fire" (or any forms), if any, in the name
```{r}
health.sal <- salary %>% filter(str_detect(Agency,"Fire"))
```

9. Make a data set called trans which contains only agencies that contain "TRANS".
```{r}
trans <- salary %>% filter(str_detect(Agency,"TRANS"))
```

10. What is/are the profession(s) of people who have "abra" in their name for Baltimore's Salaries?
```{r}
abra <- salary %>% filter(str_detect(ï..name,"Abra"))
unique(abra$JobTitle)
```


11. What is the distribution of annual salaries look like? What is the IQR?
```{r}
salary <- salary %>% mutate(AnnualSalary = AnnualSalary %>%
                                str_replace(fixed("$"),"") %>%
                                as.numeric)
quantile(salary$AnnualSalary)
```


12. Convert HireDate to the `Date` class - plot Annual Salary vs Hire Date
```{r}
salary <- salary %>% mutate(HireDate = HireDate %>%
                                 as.character %>%
                                 as.Date("%m/%d/%y"))
plot(y=salary$AnnualSalary,x=salary$HireDate)
```

13. Plot annual salary versus hire date.	Hint: first convert to numeric and date respectively

14. Create a smaller dataset that only includes the Police Department, Fire Department and Sheriff's Office. How many employees are in this new dataset?
