---
title: "DataCleaning"
author: "Angie Boysen"
date: "July 12, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Data Cleaning

Missing data with logicals
```{r}
x <- c(1,2,3,NA,7,1)
y <- 1:6
(x == 0 | x==2) & !is.na(x)
x %in% c(0,2)
```

exploring data:
table will tabulate how many things there are in each catigory - doesn't show NAs unless you ask it to
```{r table}
x
table(x)
table(x, useNA = "ifany")
table(x, useNA = "always")

tab <- table(c(0,1,2,3,2,3,3,2,2,3), c(0,1,2,3,2,3,3,4,4,3),dnn=c("x","y"))
tab
```

Recoding to missing
```{r recoding}
ages <- c(23,21,50,44,-999,12)
range(ages)
ages[ages==-999] <- NA
range(ages,na.rm = T)
```

recode missing to something
```{r recode2}
ages[is.na(ages)] <- -999
ages
```

Playing with Salary FY2104 data
```{r}
any(is.na(Sal$Name))

data$gender[data$gender %in% c("Male",'male','m','M')] <- "male"

table(gender)
```

String functions
- toupper(), tolower()
- str_trim() in stringr package or trimws() in base will trim whitespace
- nchar() number of characters in string
- paste()
- paste0() default is no space between

## Splitting, finding, replacing with Regular Expressions
What are regular expressions?
- ways to search
- find function on steriods:

. matches any single character
* means repeat as many (even if 0) more times the last character
? makes the last thing optional
^ matches the start of the vector ^a - starts with "a"
$ matches the end of the vector b$ - ends with "b"

```{r splitting}
x <- c("Angie.Boysen", "Angie","Kelsey", "Boysen")
str_split(x,fixed("."))
str_split(x, "y" )
```

Find functions
```{r grep}
grep("oy",x)
grep("oy",x, value = T)
str_detect(x, "oy")
str_subset(x, "oy")
str_extract(x, "oy")
```

changing character numbers into numbers
use sub() and gsub()
```{r}

```


